---
title: "Assignment 3"
author: "Kelsie Reinaltt"
date: "9/15/2018"
output:
  html_document: default
---
# Assignment 3: Chapters 11-13,15

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
library("nycflights13")
#install.packages("Lahman")
install.packages("datamodelr")
library("fueleconomy")
library("WHO")
```

#11.2.2 EXERCISES

## Question 1

#### I would use the read_delim() function where the first argument is the file name that you would like to read in and the second argument is the delimiter that the fields are separated by (in this case the "|").

## Question 2
#### Other arguments that they have in common include col_names, col_types, locale, na, quoted_na, quote, comment, trim_ws, skip, n_max, guess_max, progress.
```{r}
  union(names(formals(read_csv)), names(formals(read_tsv)))
```

## Question 3
#### The most important argument to read_fwf() is the col_positions one because it tells where this function to begin and end reading. 

## Question 4

# In order to read "x,y\n1,'a,b'" into the data frame, we will have to use read_csv(x,",", quote = "'") - this sets the delimiters as commas and single quotation marks. Additionally, you can also use read_csv(x,quote = "'") because the function already separates out based on commas. 

## Question 5 
#### First line -  you get 2 parsing failures. This is because two columns are created (a & b) but then the other inputs have 3 columns (since there are 2 other inputs, you get 2 parsing failures)
#### Second line - this sets 3 columns (a, b and c); however, in the second row, there are only two columns and in the third row, there are 4. 
#### Third line - this uses incorrect syntax to create a new row (should just be \n). In addition, there is only one thing in the third row in the first row there are two, and the thing is a "1" treated as a character when it is truly an integer. 
#### Fourth line - all rows are made into strings because of the first row
#### Fifth line - the values should be separated by a comma but are instead separated by a semicolon.
```{r}
read_csv("a,b\n1,2,3\n4,5,6")
read_csv("a,b,c\n1,2\n1,2,3,4")
read_csv("a,b\n\"1")
read_csv("a,b\n1,2\na,b")
read_csv("a;b\n1;3")
```

# 11.3.5

## Question 1
#### Arguments to locale() include date_names, date_format, time_format, decimal_mark, grouping_mark, tz, encoding, asciify. The importance of the arguments depends on what you are trying to do within the locale function (e.g. if you want to change the timezone, you would use the tz argument).

## Question 2
#### Locale gives you an error if you set decimal mark and grouping mark to the same thing, as shown below. 
#### When you set the decimal_mark to "," the grouping_mark is set to "." When you set the grouping mark to ".", the decimal mark is "."
```{r}
locale(decimal_mark = ".", group_mark = ".")
```
```{r}
locale(decimal_mark = ",")
```

## Question 3
#### According to the locale() documentation, the date_format() & time_format() provide the default date and time formats. The time format isn't currently used for anything, but the date format is used when guessing column types. The default date format is %Y-%m-%d. In the readr documentation, they have included an exammple: 
```{r}
parse_date("1 janvier 2015", "%d %B %Y", locale = locale("fr"))
```

## Question 4
#### In India, the locale is written as day, month, year (as opposed to the default which is month, day, year)

```{r}
india_locale <- locale(date_format = "%d/%m/%Y")

```

## Question 5
#### read_csv() uses a comma as its delimiter while read_csv2() uses a semicolon as its delimiter. 

## Question 6
#### Some of the most common encodings used in Europe include IEC 8859. In Asia, different encodings are used like GB 18030 (primarily in China) and EUC (used primarily in Japan). 


## Question 7
```{r}
d1 <- "January 1, 2010"
d2 <- "2015-Mar-07"
d3 <- "06-Jun-2017"
d4 <- c("August 19 (2015)", "July 1 (2015)")
d5 <- "12/30/14" # Dec 30, 2014
t1 <- "1705"
t2 <- "11:15:10.12 PM"

read_csv("a,b\n1,2,3\n4,5,6")
read_csv("a,b,c\n1,2\n1,2,3,4")
read_csv("a,b\n\"1")
read_csv("a,b\n1,2\na,b")
read_csv("a;b\n1;3")

parse_date(d1, "%b, %d, %y")
parse_date(d2, "%Y-%b-%d")
parse_date(d3,"%d-%b-%Y")
parse_date(d4,"%B %d (%Y)")
parse_date(d5, "%m/%d/%y")
parse_time(t1,"%H%M")
parse_time(t2, "%H:%M:%OS %p")

```



# 12.2.1
## Question 1
#### In the sample tables, there are different variables that each have observations within each column (so each variable has a column and each row consists of observations of different variables for a specific country). 
#### Table 1 - variables include country, year, cases, population. Observations are in the row and include the year, number of cases and population of each country in that year. 
#### Table 2 - variables include country, year, type and count. Observations are organized into rows and include the cases and count for each year and country
#### Table 3 - variables include country, year and rate - observations are organized into rows with the corresponding countries and years.
#### Table 4a - variables include country and two years. Observations include the country and the number of cases for that country in the two years in this case 1999 & 2000. 
#### Table 4b - variables include country and two years; observations are organized to reflect the population in that country/year. 

## Question 2
#### Because no table has both cases and population, they both required extra steps in order to calculate the cases per capita. Thus, they were both relatively similar in terms of difficulty working with. It would be easiest to work with a table that has both cases and population so that we could just use one mutate function to divide and get the correct ratio. 
```{r}
# Extract the number of TB cases per country per year.

tb_cases <- filter(table2, type == "cases") %>%
  arrange(country, year)

# Extract the matching population per country per year.

matching_population <- filter(table2, type == "population") %>%
  rename(population = count) %>%
  arrange(country, year)

# Divide cases by population, and multiply by 10000.

edited_dataframe = mutate(tb_cases, population = matching_population$population)

with_casesperpopulation = mutate(edited_dataframe, cases_per_population = count/population*10000)

select(with_casesperpopulation, country, year, cases_per_population)

# Store back in the appropriate place.
with_casesperpopulation <- with_casesperpopulation %>%
  mutate(type = "cases_per_population") %>%
  rename(count = cases_per_population)

bind_rows(table2, with_casesperpopulation) %>%
  arrange(country, year, type, count) %>%
  select(country, year, type, count)

```

```{r}
table4c <-
  tibble(country = table4a$country,
         `1999` = table4a[["1999"]] / table4b[["1999"]] * 10000,
       `2000` = table4a[["2000"]] / table4b[["2000"]] * 10000)
table4c
```

## Question 3
```{r}
table2 %>%
  filter(type == "cases") %>%
  ggplot(aes(year, count)) +
  geom_line(aes(group = country)) +
  geom_point(aes(color = country)) +
  scale_x_continuous(breaks = unique(table2$year)) +
  ylab("cases")
```

# 12.3.3

## Question 1
#### Both spread() and gather() have a convert argument. What does it do?
####convert (if true) will converty the values input to gather/spread into the appropriate type. 
#### gather() and spread() are not symmetrical because they treat the original column type differently - gather ignores the column type and standardizes everything into the same type for one uniform vecotr whereas spread() must know the type of each column
```{r}
stocks <- tibble(
  year   = c(2015, 2015, 2016, 2016),
  half  = c(   1,    2,     1,    2),
  return = c(1.88, 0.59, 0.92, 0.17)
)
stocks %>% 
  spread(year, return) %>% 
  gather("year", "return", `2015`:`2016`)
# (Hint: look at the variable types and think about column names.)
```

## Question 2
#### The code fails because 1999 and 2000 are not in quotes - in the current version of the code, 1999 and 2000 are treated as row column numbers which is not the intent of the code (rather, we are looking for years 1999 & 2000). To remedy this, we should put 1999 and 2000 in quotes. 

```{r}
table4a %>% 
gather(1999, 2000, key = "year", value = "cases")
```

##  Question 3
#### This fails because there are two rows with Phillip Woods' age. To fix this, we could add a new column that specifies the observation so that we could have two rows with Phillip Woods' age; however, the first would be observation 1 and the second would be observation 2. 
```{r}
people <- tribble(
  ~name,             ~key,    ~value,
  #-----------------|--------|------
  "Phillip Woods",   "age",       45,
  "Phillip Woods",   "height",   186,
  "Phillip Woods",   "age",       50,
  "Jessica Cordero", "age",       37,
  "Jessica Cordero", "height",   156
)
```

## Question 4
#### In order to tidy the tibble below , we need to gather it. The variables are pregnant (which can have the observation of either yes or no), sex (which can be either male or female), and count (which counts the number of instances where pregnancies occur in either males or females). 

```{r}
preg <- tribble(
  ~pregnant, ~male, ~female,
  "yes",     NA,    10,
  "no",      20,    12
)
tidied_preg <- preg %>%
  gather(male, female, key = "sex", value = "count", na.rm = TRUE)
tidied_preg

```

### 12.4.3 EXERCISES

## Question 1
#### The extra argument controls what happens when there are too many pieces (if sep is a character vector) -- there are 3 options: warn (emit a warning and drop extra values), drop (drop any extra values without warning), merge (only split at most length(into) times). Again  if sep is a character vector, fill controls what happens if there are NOT ENOUGH pieces -- there are 3 options: warn (emits a warning and fills from right, right - fills with missing values on the right), left - fills with missing values on the left. The default value in both cases is warn. 
#### experimenting done in console and agrees with above description. 
```{r}
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"))

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"))
```

## Question 2
#### remove() removes the old variable - you would set this to false if you want to create a new variable but keep the old one. 

## Question 3
#### Separate turns a single character column into multiple columns, given either regular expression or a vector of character column into multiple columns. Extract turns each group into a new column given a regular expression with capturing groups. Both of these functions turn one column into many columns - there are many ways to do this; however, unite turns multiple columns into one, which is relatively simple.


# 12.5.1 EXERCISES

## Question 1
#### Spread() allows the user to choose a specific value to replace all missing NA values. However, complete() specifies a list to replace NA values that has specific values for different variables. A similarity includes that both replacec both implicit and explicit missing values. 

## Question 2
#### Direction tells you the direction in which to fill missing values (can be down - the default - or up)


# 12.6.1 EXERCISES

## Question 1
#### In this data NA is assigned to values when there wasn't a recording for that case in a particular year. A 0 is assigned if the data was recorded but there were no cases. Since there are no implicit missing values it is okay to drop NA. 
#### A proper analysis would not exclude the missing values because it still counts as information - it is the presence of an absence. Therefore it is reasonable, but for appropriate descriptive statistics it is important to report the number of missing values.
```{r}
who %>% 
        count(country)

who %>% 
        count(country, iso2, iso3)
```

## Question 2
#### If you neglect the mutate() step, R will not know how to separate the values stored as newrel and will end up displaying a warning indicating that therethere are "too few values".

## Question 3
#### When adding the additional variables (iso2, iso3), they output the abbreviation for a country. This is redundant because we are mentioning the name of the country three times.
```{r}
library("WHO")
#select(who3, country, iso2, iso3) %>%
  #distinct() %>%
  #group_by(country) %>%
  #filter(n() > 1)
```

## Question 4

```{r}
  #who5 %>%
  #group_by(country, year, sex) %>%
  #filter(year > 1995) %>%
  #summarise(cases = sum(cases)) %>%
  #unite(country_sex, country, sex, remove = FALSE) %>%
  #ggplot(aes(x = year, y = cases, group = country_sex, colour = sex)) +
  #geom_line()
```

# 13.2.1 EXERCISES 

## Question 1

#### flights table: origin and dest
#### airports table: longitude and latitude variables
#### join flights with airports twice. The first join adds the location of the origin airport (origin). The second join adds the location of destination airport (dest).

## Question 2
#### faa in airports is matched with origin in weather.

## Question 3
#### The variables in each dataset (year/month/day/hour in both, and then origin in weather & dest in flight) would be matched.


## Question 4
#### In order to represent these special dates, I would create a dataset that has dates - the primary key in this case would be the date which would match to year/month/day and also the number of flights on that special day.

# 13.3.1 EXERCISES

## Question 1
```{r}
library("nycflights13")
flights %>%
  arrange(year, month, day, sched_dep_time, carrier, flight) %>%
  mutate(flight_id = row_number()) %>%
  glimpse()
```

## Question 2

#### 2.1 Lahman::Batting,
#### Keys are playerID, yearID, stint
#### 2.2 babynames::babynames
#### Keys are year, sex, name
#### 2.3 nasaweather::atmos
#### Keys are lat, long, year, month
#### 2.4 fueleconomy::vehicles
#### Keys are id
#### 2.5 ggplot2::diamonds
#### No keys 
#### (You might need to install some packages and read some documentation.)

## Question 3
#### How would you characterise the relationship between the Batting, Pitching, and Fielding tables?
#### All of these data sets have a 1 to 1 relationship to each other and their primary keys are playerID, yearID and stint variables. 
```{r}
library("Lahman")
library("datamodelr")
dm1 <- dm_from_data_frames(list(Batting = Lahman::Batting,
                                Master = Lahman::Master,
                                Salaries = Lahman::Salaries)) %>%
  dm_set_key("Batting", c("playerID", "yearID", "stint")) %>%
  dm_set_key("Master", "playerID") %>%
  dm_set_key("Salaries", c("yearID", "teamID", "playerID")) %>%
  dm_add_references(
    Batting$playerID == Master$playerID,
    Salaries$playerID == Master$playerID
  )

dm_create_graph(dm1, rankdir = "LR", columnArrows = TRUE)
dm2 <- dm_from_data_frames(list(Master = Lahman::Master,
                                Managers = Lahman::Managers,
                                AwardsManagers = Lahman::AwardsManagers)) %>%
  dm_set_key("Master", "playerID") %>%
  dm_set_key("Managers", c("yearID", "teamID", "inseason")) %>%
  dm_set_key("AwardsManagers", c("playerID", "awardID", "yearID")) %>%
  dm_add_references(
    Managers$playerID == Master$playerID,
    AwardsManagers$playerID == Master$playerID
  )

dm_create_graph(dm2, rankdir = "LR", columnArrows = TRUE)
```


# 13.4.6 EXERCISES 
#### Question 1
```{r}
avg_dest_delays <-
  flights %>%
  group_by(dest) %>%
  # arrival delay NA's are cancelled flights
  summarise(delay = mean(arr_delay, na.rm = TRUE)) %>%
  inner_join(airports, by = c(dest = "faa"))

avg_dest_delays %>%
  ggplot(aes(lon, lat, colour = delay)) +
    borders("state") +
    geom_point() +
    coord_quickmap()
```

#### (Don't worry if you don't understand what semi_join() does - you'll learn about it next.)
#### You might want to use the size or colour of the points to display the average delay for each airport.

## Question 2
```{r}
airport_locations <- airports %>%
  select(faa, lat, lon)

flights %>%
    select(year:day, hour, origin, dest) %>%
  left_join(
    airport_locations,
    by = c("origin" = "faa")
  ) %>%
  left_join(
    airport_locations,
    by = c("dest" = "faa")
  )
```

## Question 3
```{r}
plane_ages <-
  planes %>%
  mutate(age = 2013 - year) %>%
  select(tailnum, age)

flights %>%
  inner_join(plane_ages, by = "tailnum") %>%
  group_by(age) %>%
  filter(!is.na(dep_delay)) %>%
  summarise(delay = mean(dep_delay)) %>%
  ggplot(aes(x = age, y = delay)) +
  geom_point() +
  geom_line()
```

## Question 4
#### As shown in the graph below, increase in precipitation generally causes an increase in delay; however, the correlation is not extremely strong.
```{r}
flight_weather <-
  flights %>%
  inner_join(weather, by = c("origin" = "origin",
                            "year" = "year",
                            "month" = "month",
                            "day" = "day",
                            "hour" = "hour"))

flight_weather %>%
  group_by(precip) %>%
  summarise(delay = mean(dep_delay, na.rm = TRUE)) %>%
  ggplot(aes(x = precip, y = delay)) +
    geom_line() + geom_point()
```

## Question 5
#### According to Google, on June 13 2013, there was a large series of derechos storms especially in Southeastern US. 
```{r}
library(viridis)
flights %>%
  filter(year == 2013, month == 6, day == 13) %>%
  group_by(dest) %>%
  summarise(delay = mean(arr_delay, na.rm = TRUE)) %>%
  inner_join(airports, by = c("dest" = "faa")) %>%
  ggplot(aes(y = lat, x = lon, size = delay, colour = delay)) +
  borders("state") +
  geom_point() +
  coord_quickmap() +
  scale_colour_viridis()
```


# 13.5.1 EXERCISES
## Question 1
####When flights do have a missing tailnum, it generally means that their carrier does not report it - for example, MQ and AA (Envoy and American Airlines) do not report tailnums.
```{r}
flights %>%
  anti_join(planes, by = "tailnum") %>%
  count(carrier, sort = TRUE)
```

## Question 2
```{r}
planes_gt100 <-
  filter(flights) %>%
  group_by(tailnum) %>%
  count() %>%
  filter(n > 100)

flights %>%
  semi_join(planes_gt100, by = "tailnum")
```

## Question 3
```{r}
glimpse(fueleconomy::vehicles)
glimpse(fueleconomy::common)

fueleconomy::vehicles %>%
  semi_join(fueleconomy::common, by = c("make", "model"))
```

## Question 4
```{r}
flights %>%
  group_by(year, month, day) %>%
  summarise(total_24 = sum(dep_delay, na.rm = TRUE)+ sum(arr_delay, na.rm = TRUE)) %>%
  mutate(total_48 = total_24 + lag(total_24)) %>%
  arrange(desc(total_48))
```


## Question 5
# The first tells you the flights that go to destinations with an airport that is not in the FAA list of destinations. The second tells you US airports that don't have a flight in the dataset (i.e. no flights arrived to that location). 

## Question 6
```{r}
multi_carrier_planes <-
  flights %>%
  filter(!is.na(tailnum)) %>%
  count(tailnum, carrier) %>%
  count(tailnum) %>%
  filter(nn > 1)
multi_carrier_planes
multi_carrier_planes <-
  flights %>%
  semi_join(multi_carrier_planes, by = "tailnum") %>%
  select(tailnum, carrier) %>%
  distinct() %>%
  arrange(tailnum)
multi_carrier_planes
carrier_transfer_tbl <-
  multi_carrier_planes %>%
  group_by(tailnum) %>%
  mutate(
    carrier_num = seq_along(tailnum),
    carrier_num = paste0("carrier_", carrier_num)
  ) %>%
  left_join(airlines, by = "carrier") %>%
  select(-carrier) %>%
  spread(carrier_num, name)
carrier_transfer_tbl
```

# 15.3.1 EXERCISES 
## Question 1
```{r}
rincome_plot <-
  gss_cat %>%
  ggplot(aes(rincome)) +
  geom_bar()
rincome_plot
```
# The labels on the horizontal axis are too close together so it is impossible to read - in order to fix this, we could flip the graph so that these labels are written horizontally and taking up less space on the vertical axis as cocmpared to what they would do for the horizontal axis. 

## Question 2
#### The most common religion, as shown below, is Protestant and the most  common partyid is Independentt. 
```{r}
gss_cat %>%
  count(relig) %>%
  arrange(-n) %>%
  head(1)

gss_cat %>%
  count(partyid) %>%
  arrange(-n) %>%
  head(1)
```

## Question 3
#### It applies to protestant
```{r}
# finding out with a table
gss_cat %>%
  filter(!denom %in% c("No answer", "Other", "Don't know", "Not applicable",
                       "No denomination")) %>%
  count(relig)
```

```{r}
# finding out with a visualization
gss_cat %>%
  count(relig, denom) %>%
  ggplot(aes(x = relig, y = denom, size = n)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90))
```

# 15.4.1 EXERCISES

## Question 1
#### Although this depends on the purpose, i.e. what you are trying to use the summary for, you may want to use the median instead. The mean is affected by skewing of the data, so to get a better picture of the most realistic "normal" of the data, the median may be more accurate. However, the mean (unlike the median) does not ignore outliers, which are valid parts of the data set, so the mean can sometimes be more useful as well to get a more holistic idea of ALL data. 

## Question 2
#### The factors include:  marital, race, rincome, partyid, relig, denom.
#### As shown in the code below, principled variables include:  marital, race, and denom & arbitrary variables include rincome (real income), relig, and denom.
```{r}
levels(gss_cat[["marital"]])
gss_cat %>%
  ggplot(aes(x = marital)) +
  geom_bar()

levels(gss_cat$race)
gss_cat %>%
  ggplot(aes(race)) +
  geom_bar(drop = FALSE)

levels(gss_cat$rincome)

levels(gss_cat$relig)
gss_cat %>%
  ggplot(aes(relig)) +
  geom_bar() +
  coord_flip()

levels(gss_cat$denom)

levels(gss_cat$partyid)
```

## Question 3
#### Moving it gave it a value of 1, thereby moving it to the bottom of the plot. 

# 15.5.1 EXERCISES
## Question 1
#### A graph showing the general trend is below. None of the trends are extremely significant, but Republican identification has decreased overall, Democrat identification has remained relatively stable and independent/other has been on a slight uptrend overall. 

```{r}
gss_cat %>%
  mutate(partyid =
           fct_collapse(partyid,
                        other = c("No answer", "Don't know", "Other party"),
                        rep = c("Strong republican", "Not str republican"),
                        ind = c("Ind,near rep", "Independent", "Ind,near dem"),
                        dem = c("Not str democrat", "Strong democrat"))) %>%
  count(year, partyid)  %>%
  group_by(year) %>%
  mutate(p = n / sum(n)) %>%
  ggplot(aes(x = year, y = p,
             colour = fct_reorder2(partyid, year, p))) +
  geom_point() +
  geom_line() +
  labs(colour = "Party ID.")
```

## Question 2
```{r}
library("stringr")
gss_cat %>%
  mutate(rincome =
           fct_collapse(
             rincome,
             `Unknown` = c("No answer", "Don't know", "Refused", "Not applicable"),
             `Lt $5000` = c("Lt $1000", str_c("$", c("1000", "3000", "4000"),
                                              " to ", c("2999", "3999", "4999"))),
             `$5000 to 10000` = str_c("$", c("5000", "6000", "7000", "8000"),
                                      " to ", c("5999", "6999", "7999", "9999"))
           )) %>%
  ggplot(aes(x = rincome)) +
  geom_bar() +
  coord_flip()
```
